---
title: "Lab 09 - Grading the professor, Pt. 1"
author: "Lindsey Wilson"
date: "4/24/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r load-data}
evals <- evals
```

### Exercise 1

The distribution of `score` is visualized below:

```{r score-dist}
ggplot(evals,
       aes(x = score)) + 
  geom_histogram(stat = "count",  fill = "Grey") +
  labs(title = "Professor Evaluation Scores",
       subtitle = "Average evaluation represented by solid black line") + 
  geom_vline(xintercept = mean(evals$score), color = "Black")
```
It looks like we have some pretty substantial negative skew, which is what I expected to see. Students don't really give negative course/professor evals unless they had an extremely negative experience in the class. 

### Exercise 2

Next let's visualize the relationship between average evaluation score and average beauty rating:

```{r score-bty_avg-correlation, message = FALSE}
ggplot(evals,
       aes( x = bty_avg,
            y = score)) +
  geom_point() + 
  geom_smooth() +
  labs(title = "Average Professor Evaluation over Average Beauty Rating")

cor(evals$bty_avg, evals$score)
```

It looks like there's a weak positive correlation between average beauty rating and average professor evaluation. More attractive professors get better ratings, but the effect isn't huge.

### Exercise 3

Let's replot the above data using `geom_jitter()` instead of `geom_point()`:

```{r score-bty_avg-jitter, message = FALSE}
ggplot(evals,
       aes( x = bty_avg,
            y = score)) +
  geom_jitter() + 
  geom_smooth() +
  labs(title = "Average Professor Evaluation over Average Beauty Rating")
```

This is likely a better representation of the relationship between beauty and professor evals, because it adds a bit of random variation that wasn't present with `geom_point()`. Individuals only had a certain number of discrete options (1-5) to indicate their ratings of the professor, so the averages of those ratings were constrained to only be able to take on certain values. `geom_jitter()` fixes that problem by adding randomness to the average points in the dataset. That's why the points don't fall into neat rows and columns like they did in Exercise 2.



## 
For Exercise 12, relevel() function can be helpful! 
